---
title: "mushrooms"
author: "Julien"
date: "`r  format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r ,include=FALSE}
library(readr)
library(moments)
library(vip)
library(plyr)
library(dplyr)
library(corrplot)
library(scales)
library(Rmisc)
library(ggrepel)
library(psych)
library(ggthemes)
library(tidyr)
library(caret)
library(ggplot2)
library(pROC)
library(randomForest)
library(gbm)
library(xgboost)
library(gridExtra)
library(grid)

theme_set(theme_minimal())
```

```{r pressure, echo=FALSE}
all <- read_csv(".../Documents/Kaggle/mushrooms/datasets_478_974_mushrooms.csv")
#cleaning colnames
colnames(all) <- gsub("-","",colnames(all))
dim(all)
```

```{r}
# NA fixing: Reading description
all$gillattachment[is.na(all$gillattachment)] <- "TRUE"
```

```{r , echo=FALSE}
head(all)
```

# EDA {.tabset }

## ggplot loop function

```{r }
colNames <- colnames(all)
plot_list <- list()

for(i in colNames){
plt <- ggplot(all, aes_string(x=i, fill = "class")) +
  geom_bar(stat="count",position=position_dodge()) +
   labs(
      title= i) +
   scale_x_discrete(name="")+
    theme(axis.title.y=element_blank(),
        legend.position="none")

plot_list[[i]] <- plt
}
```

## Cap

```{r, echo=FALSE}
p1 <- plot_list$capshape
p2 <- plot_list$capsurface
p3 <- plot_list$capcolor

grobs <- grobTree(
  gp = gpar(fontsize = 14, fontface = "bold"),
  textGrob(label = "Count musrooms by classes '",
           name = "title1",
           x = unit(0.2, "lines"),
           hjust = 0, vjust = 0),
  textGrob(label = "edible",
           name = "title2",
           x = grobWidth("title1") + unit(0.2, "lines"),
           hjust = 0, vjust = 0,
           gp = gpar(col = "#F8766D")),
  textGrob(label = "' and '",
           name = "title3",
           x = grobWidth("title1") + grobWidth("title2") + unit(0.2, "lines"),
           hjust = 0, vjust = 0),
  textGrob(label = "poisonous", name = "title5",
           x = grobWidth("title1") + grobWidth("title2") + grobWidth("title3") + unit(0.2, "lines"),
           hjust = 0, vjust = 0, gp = gpar(col = "#00BFC4")),
  textGrob(label = "'.", name = "title6",
           x = grobWidth("title1") + grobWidth("title2") + grobWidth("title3") + grobWidth("title5") + unit(0.2, "lines"),
           hjust = 0, vjust = 0)
)


gg <- arrangeGrob(grobs = list(p1, p2, p3),
                  col= 3, top=grobs, padding = unit(2.6, "line"))
grid.newpage()
grid.draw(gg)
```

## Gill

```{r, echo=FALSE}
p1 <- plot_list$gillattachment
p2 <- plot_list$gillspacing
p3 <- plot_list$gillsize
p4 <- plot_list$gillcolor

grid.arrange(grobs = list(p1, p2, p3,p4),
             col= 2, top=grobs, padding = unit(2.6, "line"))
```

## stalk

```{r, echo=FALSE}
p1 <- plot_list$stalksurfaceabovering
p2 <- plot_list$stalksurfacebelowring
p3 <- plot_list$stalkcolorabovering
p4 <- plot_list$stalkcolorbelowring

grid.arrange(grobs = list(p1, p2, p3,p4),
             col= 2, top=grobs, padding = unit(2.6, "line"))
```

## veil

```{r, echo=FALSE}
p1 <- plot_list$veiltype
p2 <- plot_list$veilcolor

grid.arrange(grobs = list(p1, p2),
             col= 2, top=grobs, padding = unit(2.6, "line"))
```

## ring

```{r, echo=FALSE}
p1 <- plot_list$ringnumber
p2 <- plot_list$ringtype

grid.arrange(grobs = list(p1, p2),
             col= 2, top=grobs, padding = unit(2.6, "line"))
```

## Other variabels

```{r, echo=FALSE}
p1 <- plot_list$bruises
p2 <- plot_list$odor
p3 <- plot_list$sporeprintcolor
p4 <- plot_list$population
p5 <- plot_list$habitat

grid.arrange(grobs = list(p1, p2, p3, p4, p5),
             col= 2, top=grobs, padding = unit(2.6, "line"))
```

# Model {.tabset }

```{r, echo=FALSE}
raw.data <- all
raw.data$gillattachment <- NULL
raw.data$veiltype <- NULL

df <- data.frame(sapply(raw.data, function (x) as.numeric(as.factor(x))))
df$class <- df$class - 1
nrows <- nrow(df)
set.seed(314)
indexT <- sample(1:nrow(df), 0.7 * nrows)
trainset = df[indexT,]
testset =   df[-indexT,]
n <- names(trainset)
```

## Random Forest

```{r, echo=FALSE}
rf.form <- as.formula(paste("class ~", paste(n[!n %in% "class"], collapse = " + ")))
rftrainset <- randomForest(rf.form,trainset,ntree=100,importance=T)
```

```{r, echo=FALSE}
vipRf <-vip(rftrainset, num_features = 20, fill = "#a3a500")
```

```{r, echo=FALSE}
testset$predicted <- round(predict(rftrainset ,testset),0)
testset_rf <- testset$predicted;
```

```{r, echo=FALSE}
plotConfusionMatrix <- function(testset, sSubtitle,lowCol, higCol) {
    tst <- data.frame(testset$predicted, testset$class)
    opts <- c("Predicted", "True")
    names(tst) <- opts
    lowCol <- lowCol
    higCol <- higCol
      
    cf <- plyr::count(tst)
    cf[opts][cf[opts]==0] <- "Poisonous"
    cf[opts][cf[opts]==1] <- "Edible"

    ggplot(data =  cf, mapping = aes(x = True, y = Predicted)) +
      labs(title = "Confusion matrix", subtitle = sSubtitle) +
      geom_tile(aes(fill = freq), colour = "grey") +
      geom_text(aes(label = sprintf("%1.0f", freq)), vjust = 1) +
      scale_fill_gradient(low = lowCol , high = higCol ) +
      theme_bw() + theme(legend.position = "none")
}

plotConfusionMatrix2 <- function(testset,lowCol, higCol) {
    tst <- data.frame(testset$predicted, testset$class)
    opts <- c("Predicted", "True")
    names(tst) <- opts
    lowCol <- lowCol
    higCol <- higCol
      
    cf <- plyr::count(tst)
    cf[opts][cf[opts]==0] <- "Poisonous"
    cf[opts][cf[opts]==1] <- "Edible"

    ggplot(data =  cf, mapping = aes(x = True, y = Predicted)) +
      geom_tile(aes(fill = freq), colour = "grey") +
      geom_text(aes(label = sprintf("%1.0f", freq)), vjust = 1) +
      scale_fill_gradient(low = lowCol , high = higCol ) +
      theme_bw() + theme(legend.position = "none")
}
```

```{r fig.width=4, fig.height=4, echo=FALSE}
plotConfusionMatrix(testset,"Prediction using RandomForest with 100 trees","#a3a500","#5d5e02")
matRf <- plotConfusionMatrix2(testset,"#a3a500","#5d5e02")
```

## Gradient Boosting Machine model

```{r, echo=FALSE}
n <- names(trainset)
gbm.form <- as.formula(paste("class ~", paste(n[!n %in% "class"], collapse = " + ")))
```

```{r, echo=FALSE}
gbmFormula = gbm(formula = gbm.form,
               distribution = "bernoulli",
               data = trainset,
               n.trees = 500,
               shrinkage = .1,
               n.minobsinnode = 15,
               cv.folds = 5,
               n.cores = 1)
```

```{r include=FALSE}
optimalTreeNumberPredictionCV <- gbm.perf(gbmFormula)
```

```{r, echo=FALSE}
gbmTest = predict(object = gbmFormula,
                           newdata = testset,
                           n.trees = optimalTreeNumberPredictionCV,
                           type = "response")
testset$predicted <- round(gbmTest,0)
testset_gbm <- testset$predicted
```

```{r, echo=FALSE}
vip(gbmFormula, num_features = 20, fill = "#00bf7c") + ggtitle("Variable importance of xgb")
vipGb <- vip(gbmFormula, num_features = 20, fill = "#00bf7c")
```

```{r fig.width=4, fig.height=4}
plotConfusionMatrix(testset,"Prediction using RandomForest with 100 trees", "#01784e","#00bf7c")
matGb <- plotConfusionMatrix2(testset, "#01784e","#00bf7c")
```

## XGBoost model

```{r echo=FALSE}
dmTrain <- xgb.DMatrix(as.matrix(trainset %>% select(-class)), label = trainset$class)
dmTest <- xgb.DMatrix(as.matrix(testset %>% select(-class,-predicted)), label = testset$class)
```

```{r echo=FALSE}
params <- list(
  "objective" = "binary:logistic",
  "eval_metric" = "auc",
  "eta" = 0.012,
  "subsample" = 0.8,
  "max_depth" = 8,
  "colsample_bytree" = 0.9,
  "min_child_weight"= 5
)
```

```{r echo=FALSE}
model_xgb.cv <- xgb.cv(params=params,
                      data = dmTrain, 
                      maximize = TRUE,
                      nfold = 5,
                      nrounds = 5000,
                      nthread = 3,
                      early_stopping_round=100,
                      print_every_n=100)
```

```{r echo=FALSE}
model_xgb <- xgboost(params=params,
                      data = dmTrain, 
                      maximize = TRUE,
                      nrounds = 5000,
                      nthread = 3,
                      early_stopping_round=100,
                      print_every_n=100)
```

```{r echo=FALSE}
testset$predicted <- round(predict(model_xgb ,dmTest),0)
testset_xgboost <- testset$predicted
```

```{r echo=FALSE}
vip(model_xgb, num_features = 20, fill = "#e76bf3") + ggtitle("Variable importance of xgb")
vipXgb <- vip(model_xgb, num_features = 20, fill = "#e76bf3")
```

```{r fig.width=4, fig.height=4, echo=FALSE}
plotConfusionMatrix(testset,"Prediction using RandomForest with 100 trees", "#83378a","#e76bf3")
matXgb <- plotConfusionMatrix2(testset, "#83378a","#e76bf3")
```

```{r echo=FALSE}
print(sprintf("Area under curve (AUC) : %.3f",auc(testset$class, testset$predicted)))
```

# Model comparison

```{r echo=FALSE}
grobs <- grobTree(
  gp = gpar(fontsize = 12, fontface = "bold"),
  textGrob(label = "Variable importance and confusion matrix of ",
           name = "title1",
           x = unit(0.2, "lines"),
           hjust = 0, vjust = 0),
  textGrob(label = "Random Forest",
           name = "Rf",
           x = grobWidth("title1") + unit(0.2, "lines"),
           gp = gpar(col = "#a3a500"),
           hjust = 0, vjust = 0),
  textGrob(label = ", ",
           name = ",",
           x = grobWidth("title1") + grobWidth("Rf") + unit(0.2, "lines"),
           hjust = 0, vjust = 0),
  textGrob(label = "Gradient Boosting",
           name = "GB",
           x = grobWidth("title1") + grobWidth("Rf")+ grobWidth(",") + unit(0.2, "lines"),
           gp = gpar(col = "#00bf7c"),
           hjust = 0, vjust = 0),
  textGrob(label = " and ",
           name = "and",
           x = grobWidth("title1") + grobWidth("Rf")+ grobWidth(",")+ grobWidth("GB") + unit(0.2, "lines"),
           hjust = 0, vjust = 0),
  textGrob(label = "Extrem Gradient Boosting",
           name = "XGB",
           x = grobWidth("title1") + grobWidth("Rf")+ grobWidth(",")+ grobWidth("GB")+ grobWidth("and") + unit(0.2, "lines"),
           gp = gpar(col = "#e76bf3"),
           hjust = 0, vjust = 0)
  )
```

```{r fig.width=9, fig.height=6, echo=FALSE}
grid.arrange(grobs = list(vipRf, vipGb, vipXgb, matRf,matGb, matXgb), 
             ncol= 3, top=grobs, padding = unit(2.6, "line"))
```

# Conclusions

We were able to predict with very high accuracy the poisonous and edible mushrooms based on the three models used, `Random Forest, Gradient Boosting Machine (GBM) and XGBoost. For the GBM and XGBoost models we were also using cross validation. The best prediction was obtained using Random Forest model.